Acceleration of Derivative Calculations with Application to Radial Basis Function -- Finite-Differences on the Intel MIC Architecture

### 2014-01-17 acm/ieee_paper.tex
edits are done

Figure~\ref{fig:bandwidth} shows a benchmark that computes the
bandwidth achieved when reading and/or writing large arrays on both
the MIC card and the CPU architecture. The read benchmark performs a
simple sum of the array. The write benchmark sets all the elemnts of
the array to zero, while the read/write benchmarks copies the array into
another one. We also investigate the difference between a straight C++
implementation (compiled with \mbox{-O3} with proper tagging to inform the
compiler of memory alignments) and a simple implementation with
explicit use of the vector registers.

@@

To put things in perspective, we conducted similar experiments on a
classical CPU architecture with two Intel Core i7 processors, each has eight
cores that support AVX instructions. Both the compiler-generated and
the manually-written vectorization using AVX instructions reach a
(read) bandwidth of 70~GB/s. This indicates that the performance of
the classical SpMV can not overcome 17~Gflop/s. A generic
implementation of the multiplication of four vectors with four matrices on
the {\tt RBF} matrix reaches a performance of 27~Gflop/s; while a
manual implementation using AVX instructions reaches a performance of
42~Gflop/s. This confirms that our method is sound and generalizes to
architectures other than the Intel MIC architecture.

@@

In future work, we will examine the effect of larger stencil sizes and
double precision. The presented performance does not reach the predicted
peak and we believe that common techniques from SpMV such as cache
partitioning can be successfully applied. We will also integrate the
techniques presented in this paper within a fluid simulation using
RBFFD.

>>>>

Figure~\ref{fig:bandwidth} shows a benchmark that computes the
bandwidth achieved when reading and/or writing large arrays on 
the MIC card. The read benchmark performs a
simple sum of the array. The write benchmark sets all the elemnts of
the array to zero, while the read/write benchmarks copies the array into
another one. We also investigate the difference between a straight C++
implementation (compiled with \mbox{-O3} with proper tagging to inform the
compiler of memory alignments) and an implementation with
explicit use of the vector registers.

@@

\ee{The main idea of this paper, computing 16 derivatives at a time,
  can be applied on different architectures such as multiprocessor
  multicore or GPU. To illustrate this fact, we conducted similar
  experiments on a classical CPU architecture with two Intel Core i7
  (SandyBridge) processors, each has eight cores that support AVX
  instructions.}  Both the compiler-generated and the manually-written
vectorization using AVX instructions reach a (read) bandwidth of
70~GB/s. This indicates that the performance of the classical SpMV can
not overcome 17~Gflop/s. A generic implementation of the
multiplication of four vectors with four matrices on the {\tt RBF}
matrix reaches a performance of 27~Gflop/s. We also implemented an AVX
vectorized version which is less compact than the one on MIC because
of smaller vector registers and the abscence of a gather instruction
(AVX2 on Haswell processor has that feature). It reaches a
performance of 42~Gflop/s. This confirms that our method is sound and
generalizes to architectures other than the Intel MIC architecture.

@@

In future work, we will examine the effect of larger stencil sizes and
double precision. The presented performance does not reach the
predicted peak and we believe that common techniques from SpMV such as
cache partitioning can be successfully applied. We will also integrate
the techniques presented in this paper within a fluid simulation using
RBFFD. We will also look developping a similar technique on more
commonly available devices such as GPUs.
